---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

## Loading and preprocessing the data

```{r}
unzip("activity.zip")
data <- read.csv("activity.csv")
summary(data)

library(plyr)

data$d1 <- sprintf("%04d", data$interval)
data$interval_dt <- strptime(data$d1, "%H%M" )
data$date_dt <- strptime(data$date, "%Y-%m-%d")
data$d1 <- NULL

stats <- ddply(data, .(date), summarise, 
               "steps_per_day" = sum(steps, na.rm = T))

time_stats <- ddply(data, .(interval_dt), summarise,
                "avg_steps_per_int" = mean(steps, na.rm = T))

```

## What is mean total number of steps taken per day?
```{r}

hist(stats$steps_per_day, breaks=
       seq(0,max(stats$steps_per_day)+500, by=500))

mean_per_day <- mean(stats$steps_per_day, na.rm=T)
median_per_day <- median(stats$steps_per_day, na.rm=T)

```

Mean value of steps per day is `r mean_per_day`.
Median value of steps per day is `r median_per_day`.


## What is the average daily activity pattern?

```{r, fig.width=15 }

startTime <- strptime("00:00", "%H:%M")
endTime <- strptime("23:55", "%H:%M")

plot(time_stats$interval_dt, time_stats$avg_steps_per_int, 
     type="l", axes = F)
axis(side = 2)
axis.POSIXct(side = 1, las = 2, at = seq(startTime, endTime, by="hour"))

steps_max <- max(time_stats$avg_steps_per_int)
time_int_max <- time_stats$interval_dt[time_stats$avg_steps_per_int == steps_max]

time_int_max_fmt <- format(time_int_max, "%H:%M")

```

Maximum average number of steps per day is `r steps_max` for interval `r time_int_max_fmt`.


## Imputing missing values

Fixing NA values with a very simple strategy - by replacing them with mean value for this time period. 


```{r}

dataFixed <- data

for(i in 1:length(dataFixed$steps))
{
  if(is.na(dataFixed$steps[i]))
    dataFixed$steps[i] <- time_stats$avg_steps_per_int[i]
}

statsFixed <- ddply(dataFixed, .(date), summarise, 
               "steps_per_day" = sum(steps, na.rm = T))

time_statsFixed <- ddply(dataFixed, .(interval_dt), summarise,
                "avg_steps_per_int" = mean(steps, na.rm = T))

hist(statsFixed$steps_per_day, breaks=
       seq(0,max(stats$steps_per_day)+500, by=500))

mean_per_day_fixed <- mean(statsFixed$steps_per_day, na.rm=T)
median_per_day_fixed <- median(statsFixed$steps_per_day, na.rm=T)

#print(head(dataFixed))

```
Mean value of steps per day with fixed NA values is `r mean_per_day_fixed` while original mean value is `r mean_per_day`.  
Median value of steps per day with fixed NA values is `r median_per_day_fixed` while original median value is `r median_per_day`.  


## Are there differences in activity patterns between weekdays and weekends?

```{r fig.align='center', fig.width=15  }

dataFixed$day <- dataFixed$date_dt$wday

weekendsData <- dataFixed[dataFixed$day == 0 | dataFixed$day == 6, ]
weekdayData <- dataFixed[dataFixed$day != 0 & dataFixed$day != 6, ]

time_statsFixedWeekend <- ddply(weekendsData, .(interval_dt), summarise,
                "avg_steps_per_int" = mean(steps, na.rm = T))

time_statsFixedWeekday <- ddply(weekdayData, .(interval_dt), summarise,
                "avg_steps_per_int" = mean(steps, na.rm = T))

par(mfrow = c(2,1))
plot(time_statsFixedWeekend$interval_dt, time_statsFixedWeekend$avg_steps_per_int, type="l", axes = F, main="weekend")
axis(side = 2)
axis.POSIXct(side = 1, las = 2, at = seq(startTime, endTime, by="hour"))

plot(time_statsFixedWeekday$interval_dt, time_statsFixedWeekday$avg_steps_per_int, type="l", axes = F, main="weekday")
axis(side = 2)
axis.POSIXct(side = 1, las = 2, at = seq(startTime, endTime, by="hour"))

```



